---
title: "Biol432 Tutorial - 1 Markdown"
author: "Sree"
date: "`r Sys.Date()`"
output: pdf_document
---

## Load Required Libraries and Data

```{r setup, include=TRUE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(tinytex)

# Load the revised measurements.csv
data <- read.csv("measurements.csv")

# Display the first few rows of the dataset
head(data)
```

## Sorting of the data

```{r sorting of data, include=TRUE}
# Sort data by Species, Observer, and Volume
sorted_data <- data %>%
  arrange(Species, Observer, Volume)

# Display the sorted dataset
head(sorted_data)
```

## Calculating the average limb volume of species

```{r average limb volume of species}
# Create a table showing the average volume for each species
average_volume <- data %>%
  group_by(Species) %>%
  summarize(Average_Volume = mean(Volume, na.rm = TRUE))


## Create a box plot comparing limb volumes by species
```{r }
ggplot(data, aes(x = Species, y = Volume)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  theme_minimal() +
  labs(
    title = "Distribution of Limb Volume by Species",
    x = "Species",
    y = "Limb Volume")

```{r}
# Create frequency histograms for limb volume by species
ggplot(data, aes(x = Volume)) +
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "darkgreen") +
  facet_wrap(~ Species, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Frequency Histograms of Limb Volume by Species",
    x = "Limb Volume",
    y = "Frequency")











